Paper_Notes

### 2017-10-12 Paper_Notes.Rmd
Summary #4 Commit

>>>>

### Atman et al. 2014

- Altman, Naomi, and Martin Krzywinski. "Points of Significance: Sources of Variation." Nature Methods 12, no. 1 (December 30, 2014): 5-6.

The article discusses the double-edged sword that is variation in biological based experiments. Although general consensus is to minimize experimental variability, some variability needs to be maintained to allow generalization of the results. The authors argue that experimental control, randomization, blocking, and replication are the tools that allow replicable and meaningful results. Variability affects replication of experimental results (internal validity) which, in turn, affects the generalizability to a larger population (external validity). A well-designed experiment is a compromise between internal and external validity, and, with careful experimental design, the impact of biological variability can be reduced. Genotype and gender are examples of sources of variability that are under complete experimental control, while other sources of variability, such as diet, temperature, and housing, are under partial experimental control. So-called "noise" factors that cannot be controlled, or are unknown, can be handled by randomization, replication, and blocking. Randomization is employed to avoid biasing experimental results. Replication is needed to increase the precision with which a sample is characterized (e.g. standard error mean). However, the biological variability can be underestimated when measurements are an average of a large number of contributing factors. Blocking on a noise variable removes its effect by taking a difference of two measurements that share the same value of the noise, effectively isolating this variability source. Technical variability, such as reagents, measurement platforms, and personnel, also contribute to the overall variation. Steps should be taken to reduce sources of variation that are nuisance factors while maintaining the sources of variation required to assess how much effects vary in the generalizable population. Whereas the former should be minimized to optimize the power of the experiment, the latter need to be sampled and quantified so that results can be generalized while maintaining the integrity of the estimate uncertainty.
